---
title: "Matrices"
author: "Sarah McDonald"
date: "7/19/2022"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 6
    toc_float: yes
---

__Purpose:__

This script runs the VAR and BigVar models on our simulated dataset.

I show all combinations for $n = \{4, 10, 20\}$ and $t = \{100, 200, 500\}$

__Takeaways:__

- None of the BigVAR model coefficients look like we would expect. Most are 0. 

- When $n = 200$ in the VAR model, there is some strange behavior. The coefficients become quite large, and some become NA. Still, the nonmissing diagonals are very close to what we would expect. Still wanted to flag this as strange behavior.


```{r, warning = F, message=F}
library(igraph)
library(tidyverse)
library(BigVAR)
library(vars)
library(knitr)
library(kableExtra)

models <- function(n, t, rho_1_mod, rho_2_mod=0, p_num, s_alpha_input, s_epsilon_input){
  set.seed(4)
  N = n
  s_epsilon = s_epsilon_input
  s_alpha = s_alpha_input
  s_beta = 0
  
  p = p_num/N
  
  rho_1 = rho_1_mod
  rho_2 = rho_2_mod
  
  
  y_0 = rep(0, N)
  
  
  alpha = rnorm(N, mean = 0, sd = s_alpha)
  
  # beta is the same as time dimension, for each time stamp there is a time value, when we eventually turn it on
  beta = rnorm(N, mean = 0, sd = s_beta)
  epsilon = rnorm(N, mean = 0, sd = s_epsilon)
  
  
  G <- erdos.renyi.game(N, p, type=c("gnp"), directed = FALSE, loops = F) %>%
    as_adjacency_matrix(sparse = F)
  
  
  periods = t
  Y <- matrix(nrow = N, ncol = periods + 1)
  colnames(Y) <- 0:periods
  
  Y[, "0"] <- rnorm(N, mean = 0, sd = 1)
  
  for(t1 in 1:periods){
    
    t = t1 - 1
    
    y_t <- Y[, paste(t)]
    
    
    y_t1 <- alpha + beta + rho_1 * y_t + (rho_2 * G %*% y_t )+ epsilon
    
    Y[, paste(t1)] <- y_t1
    
  }
  
  identity = diag(N)
  
  comparison_matrix = ((rho_1 * identity) +( rho_2 *G))
  
  # big var
 # big_var_model=BigVAR.fit(t(Y),p = 1,
  #                 struct = "Basic",
  #                 lambda=1e-20, 
 #                  intercept=F)
  
#  big_var_mod_subset <- big_var_model[,,1]
  
  
  
  # var
  model <- VAR(t(Y), p = 1)
  
  model_matrix <-  t(as.data.frame(lapply(model$varresult, `[[`, 1)))
  
  
  
  rownames(model_matrix) <- paste("y", 1:N, sep = "")
  
 # return(
  #  list(
  #    print("rho_1:"), print(rho_1),
  #    print("VAR Matrix:"), print(format(round(model_matrix, 10), scientific = F)),
  #    print("Alpha"), print(alpha),
  #    print("Epsilon"), print(epsilon),
   
  # print("Adjacency Matrix"), print(G)
         # paste("BigVAR Matrix:"),
  #  format(round(big_var_mod_subset, 10), scientific=F)
  #)
  #)
  
  return(list(list("Rho_1:", rho_1), 
              list("Alpha: ", alpha), 
              list("Epsilon:", epsilon), 
              list("VAR Matrix:", format(round(model_matrix, 10), scientific = F)),
              list("Adjaceny Matrix:", G)))

  
}
```



## Output {.tabset}

### T = 100 {.tabset}

#### n = 4
```{r, warning = F, message=F}
N= 4
models(n = N, t = 100, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```
#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 100, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)

```
#### n = 20
```{r, warning = F, message=F}
N= 20
models(n = N, t = 100, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

### T = 200 {.tabset}

#### n = 4
```{r, warning = F, message=F}
N= 4
models(n = N, t = 200, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```
#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 200, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```
#### n = 20
```{r, warning = F, message=F}
N= 20
models(n = N, t = 200, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

### T = 500 {.tabset}

#### n = 4
```{r, warning = F, message=F}
N= 4
models(n = N, t = 500, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```
#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 500, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```
#### n = 20
```{r, warning = F, message=F}
N= 20
models(n = N, t = 500, rho_1_mod = runif(N, min = 0.7, max = 1.3), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```




## Output - Random Normal {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 200, rho_1_mod = rnorm(N), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 500, rho_1_mod = rnorm(N), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```



## Output - smaller uniform range{.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10

models(n = N, t = 200, rho_1_mod = runif(N, min = 0.95, max = 1.05), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)


```

### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 500, rho_1_mod = runif(N, min = 0.95, max = 1.05), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

## Output - smaller number range [0.7, 0.9]{.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N = 10
models(n = N, t = 200, rho_1_mod = runif(N, min = 0.7, max = 0.9), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N = 10
models(n = N, t = 500, rho_1_mod = runif(N, min = 0.7, max = 0.9), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```


## Output - big number range [-1, 1] {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 200, rho_1_mod =runif(N, min = -1, max = 1), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```

### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(n = N, t = 500, rho_1_mod =runif(N, min = -1, max = 1), rho_2_mod=0, p_num = 1, s_alpha_input = 0, s_epsilon_input = 0)
```


# new Rho_2






## Rho_2 = 1/N, p numerator= 3 {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 200, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input =0, s_epsilon_input=0)
```


### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 500, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input =0, s_epsilon_input=0)
```






## Rho_2 = 1/N, p numerator= 3, s_alpha = 0.1 {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 200, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 0.1, s_epsilon_input=0)
```


### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 500, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 0.1, s_epsilon_input=0)
```


## Rho_2 = 1/N, p numerator= 3, s_alpha = 0.1, s_epsilon = 0.1 {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 200, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 0.1, s_epsilon_input=0.1)
```


### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 500, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 0.1, s_epsilon_input=0.1)
```




## Rho_2 = 1/N, p numerator= 3, s_alpha = 1, s_epsilon = 0.1 {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 200, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 1, s_epsilon_input=0.1)
```


### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 500, rho_1_mod = runif(N, min = -1, max = 1), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 1, s_epsilon_input=0.1)
```


## rho_1 = 1 Rho_2 = 1/N, p numerator= 3, s_alpha = 1, s_epsilon = 0.1 {.tabset}

### T = 200 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 200, rho_1_mod = rep(1, N), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 1, s_epsilon_input=0.1)
```


### T = 500 {.tabset}

#### n = 10
```{r, warning = F, message=F}
N= 10
models(N, 500, rho_1_mod =rep(1, N), rho_2_mod = 1/N, p_num = 3, s_alpha_input = 1, s_epsilon_input=0.1)
```




# How big can T be given N?

We want to lower T and see if T needs to scale with N or N^2.

In the table, I show different $\rho_1$, sample size $N$, and values $t$ configurations. I count "weird" behavior: (1) number of NA columns outputed and (2) number of large coefficients, where the absolute value of the coefficient must be greater than 10 to be flagged as a large coefficients.

Below, we list all possibilities for $N$, $t$, and $\rho_1$:

$N \in \{4, 10, 20, 25\}$

$t(n) \in \{n + 2, n^2 + n +1\}$

$\rho_1 \in \{Unif[0.7, 1.3], Unif[.95, 1.05], Unif[-1, 1], N(0, 1)\}$



The blue highlighted rows indicate rows without any atypical behavior. 



For $N \geq 30$, I am finding that VAR breaks at $t(n) = N^2 + N + 1$. This is the specific error: `Error in VAR(t(Y), p = 1) : 
NAs in y.`




```{r, warning=F, message=F}
N_list <- c(4, 10, 20, 25)
p_num_list <- c(1, 3)

rho_1_list <- c("unif_0.7_1.3", "unif_0.95_1.05", "unif_-1_1", "std_normal")
t_type_list <- c("N", "N-squared")


# df of all configurations!
df <- data.frame(N = rep(N_list, each = length(t_type_list)*length(rho_1_list) * length(p_num_list)), 
                 rho_1 = rep(rep(rho_1_list, each=length(t_type_list)*length(p_num_list)), length(N_list)), 
                 p_num = rep(rep(p_num_list, each = length(t_type_list)), length(rho_1_list)*length(N_list)), 
                 t_type = rep(t_type_list, length(p_num_list)* length(N_list)* length(rho_1_list)), 
                 t = NA,
                 NA_cols = NA,
                 large_coefs = NA, rho_2_mod=0,  s_alpha_input = 0, s_epsilon_input = 0) %>%
  mutate(t = case_when(t_type== "N" ~ N + 2,
                       t_type== "N-squared" ~ N^2 + N+ 1))

for(N in N_list){
  set.seed(1)
  rho_1_list <- list(`unif_0.7_1.3`= runif(N, min = 0.7, max = 1.3), 
                     `unif_0.95_1.05` = runif(N, min = 0.95, max = 1.05), 
                     `unif_-1_1` = runif(N, min = -1, max = 1),
                     std_normal =  rnorm(N))
  
  t_list <- list(`n` = N+2, `n_squared` = N^2 + N +1)
  
  for(i in names(rho_1_list)){
    rho_1_name <- i
    rho_1_value <- rho_1_list[[i]]
    
    for(j in t_list){
      
      for(k in p_num_list){
      
      m <- models(n = N, t = j, rho_1_mod = rho_1_value, rho_2_mod=0, p_num = k, s_alpha_input = 0, s_epsilon_input = 0)[[4]][[2]]
      class(m) <- "numeric"
      
      # number of large coefficients
      df[df$N == N & df$rho_1 == rho_1_name & df$t == j & df$p_num == k, "large_coefs"] <- sum(abs(m)> 10, na.rm = T)
     
      # number of NA columns
      df[df$N == N & df$rho_1 == rho_1_name & df$t == j & df$p_num == k, "NA_cols"] <- length(colnames(m)[colSums(is.na(m)) > 0])
      }
    }
  }
}


kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$NA_cols == 0 & df$large_coefs == 0), background = "lightblue") %>%
  scroll_box(height = "500px")

```











